---
title: "Modelo predictivo salarios NBA"
author: "Mario Arteche Miranda"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---
# Descripción de las variables

**Player**: Name and surname.

**Salary**: Players salary.

**NBA_Country** : Players country.

**NBA_DraftNumber**: Player number.

**Age**: Players age.

**Tm**: Players actual team.

**G** : Number of played matches.

**MP** : Time in minutes played.

**PER**: Player Efficiency Rating A measure of per-minute production standardized such that the league average is 15.

**TS.** : True Shooting Percentage A measure of shooting efficiency that takes into account 2-point field goals, 3-point field goals, and free throws.

**X3PAr** : 3-Point Attempt Rate Percentage of FG Attempts from 3-Point Range.

**FTr** : Free Throw Attempt Rate Number of FT Attempts Per FG Attempt.

**ORB.**:  Offensive Rebound Percentage An estimate of the percentage of available offensive rebounds a player grabbed while he was on the floor.

**DRB. **: Defensive Rebound Percentage An estimate of the percentage of available defensive rebounds a player grabbed while he was on the floor.

**TRB.** : Total Rebound Percentage An estimate of the percentage of available rebounds a player grabbed while he was on the floor.

**AST.** : Assist Percentage An estimate of the percentage of teammate field goals a player assisted while he was on the floor.

**STL.** : Steal Percentage An estimate of the percentage of opponent possessions that end with a steal by the player while he was on the floor.

**BLK.** : Block Percentage An estimate of the percentage of opponent two-point field goal attempts blocked by the player while he was on the floor. 

**TOV.** : Turnover Percentage An estimate of turnovers committed per 100 plays.

**USG.** : Usage Percentage An estimate of the percentage of team plays used by a player while he was on the floor.

**OWS** : Offensive Win Shares An estimate of the number of wins contributed by a player due to his offense.

**DWS** : Defensive Win Shares An estimate of the number of wins contributed by a player due to his defense.

**WS** : Win Shares An estimate of the number of wins contributed by a player.

**WS.48** : Win Shares Per 48 Minutes An estimate of the number of wins contributed by a player per 48 minutes (league average is approximately .100)

**OBPM** : Offensive Box Plus/Minus A box score estimate of the offensive points per 100 possessions a player contributed above a league-average player, translated to an average team.

**DBPM** : Defensive Box Plus/Minus A box score estimate of the defensive points per 100 possessions a player contributed above a league-average player, translated to an average team.

**BPM** : Box Plus/Minus A box score estimate of the points per 100 possessions a player contributed above a league-average player, translated to an average team.

**VORP** : Value over Replacement Player A box score estimate of the points per 100 TEAM possessions that a player contributed above a replacement-level (-2.0) player, translated to an average team and prorated to an 82-game season.


# Cargamos las librerias a emplear lectura y procesado de los datos
```{r tidy=TRUE,message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(readr)
library(gvlma)
library(MASS)
library(car)
library(formatR)
library(tidyverse)
```

# lectura y procesado de los datos
```{r tidy=TRUE,message=FALSE, warning=FALSE, include=FALSE}
mData <- read.csv('nba.csv')
attach(mData)
head(mData)
names(mData)
```

* Tratamiento de los datos

## Valores unicos del data set.
```{r echo=FALSE}
distintos<-distinct(mData)
head(distintos)
```

## Valores unicos de la columna Player
```{r echo=FALSE}
distpla<-distinct(mData,mData$Player)
head(distpla)
```

## Numero de jugadores repetidos y eliminacion de estos
```{r echo=FALSE}
nrow(mData[duplicated(mData$Player),])#vmos valores repetidos en jugadores
mData<-mData[!duplicated(mData$Player),]#eliminamos los valores repetidos
nrow(mData[duplicated(mData$Player),])#vemos si vuelven a estar esos duplicados
```

# Construimos nuestro modelo
### Previamente cargamos los nombres de las columnas de la base de datos para transcribirlos
```{r echo=FALSE}
regres01  <- lm(Salary ~ NBA_DraftNumber + Age + G + MP + PER + TS. + X3PAr + FTr + ORB. + DRB. + TRB. + AST. + STL. + BLK. + TOV. + USG. + OWS+ DWS + WS + WS.48 + OBPM + DBPM + BPM + VORP,data = mData)
summary(regres01)
```

## Una vez cargado y "analizado" el modelo, analizamos si cumple las 4 hipotesis de significatividad, con el modelo global
```{r echo=FALSE}
gvmodel <- gvlma(regres01) 
summary(gvmodel)#vemos que solo cumple 1 de las 4 hipotesis(heterocedasticidad, varianza constante) para ser explicativo
```

## Tras el apartado anterior, vemos que el modelo no satisface nuestras hipotesis, por lo que deberemos seleccionar aquellas variables que deberemos eliminar o modificar
```{r echo=FALSE}
stepAIC(regres01,direction = "both")
#este modelo es el que mejor se ajusta:
```

## Creamos el modelo a partir del menor de los AIC
```{r echo=FALSE}
regres02 <- lm(formula = Salary ~ NBA_DraftNumber + Age + G + MP + PER + 
               X3PAr + ORB. + TRB. + USG. + WS + OBPM, data = mData)
summary(regres02)
```

## Validacion global del segundo modelo
```{r echo=FALSE}
gvmodel2<-gvlma(regres02)
summary(gvmodel2)## Sigue cumpliendo solamente una de las 4 hipotesis (heterocedasticidad)
```

## Vamos a analizar de manera grafica cada una de las hipotesis
### Normalidad
```{r}
qqPlot(regres02, labels=row.names(states), id.method="identify",
       simulate=TRUE, main="Q-Q Plot")
```

### Kernel+hist+normalidad
```{r echo=FALSE}
residplot <- function(regres02, nbreaks=10) {
  z <- rstudent(regres02)
  hist(z, breaks=nbreaks, freq=FALSE,
       xlab="Studentized Residual",
       main="Distribution of Errors")
  rug(jitter(z), col="brown")
  curve(dnorm(x, mean=mean(z), sd=sd(z)),
        add=TRUE, col="blue", lwd=2)
  lines(density(z)$x, density(z)$y,
        col="red", lwd=2, lty=2)
  legend("topright",
         legend = c( "Normal Curve", "Kernel Density Curve"),
         lty=1:2, col=c("blue","red"), cex=.7)
}
residplot(regres02)
```

### Linealidad
```{r echo=FALSE}
crPlots(regres02)
```

### Heterocedasticidad: rechazamos la hipotesis de homocedasticidad
```{r echo=FALSE}
ncvTest(regres02) 
```

### Heterocedasticidad grafica
```{r echo=FALSE}
spreadLevelPlot(regres02) 
```

### Problemas de colinealidad
#### Vemos que lo hay, porque varios superan el **2**
```{r}
sqrt(vif(regres02)) > 2 

## Analisis de valores atipicos
```{r echo=FALSE}
outlierTest(regres02)
```

### Representacion grafica de los outliers: Parecen ser influyentes
```{r warning=FALSE, echo=FALSE}
influencePlot(regres02, id.method="identify", main="Influence Plot", 
              sub="Circle size is proportial to Cook's Distance" ) 
```

## Importancia de las variables
```{r relative_weights,echo=FALSE}

# Definimos la funcion
relweights <- function(fit,...){ 
  R <- cor(fit$model) 
  nvar <- ncol(R)
  rxx <- R[2:nvar, 2:nvar] 
  rxy <- R[2:nvar, 1] 
  svd <- eigen(rxx) 
  evec <- svd$vectors 
  ev <- svd$values
  delta <- diag(sqrt(ev))
  lambda <- evec %*% delta %*% t(evec)
  lambdasq <- lambda ^ 2 
  beta <- solve(lambda) %*% rxy 
  rsquare <- colSums(beta ^ 2) 
  rawwgt <- lambdasq %*% beta ^ 2 
  import <- (rawwgt / rsquare) * 100 
  import <- as.data.frame(import)
  row.names(import) <- names(fit$model[2:nvar])
  names(import) <- "Weights"
  import <- import[order(import),1, drop = FALSE]
  dotchart(import$Weights, labels = row.names(import), xlab = "% of R-Square", pch = 19, 
           main = "Relative Importance of Predictor variables", 
           sub = paste("Total R-Square=", round(rsquare, digits = 3)),
           ...)
return(import)
}
```

### Aplicamos la funcion
```{r echo=FALSE}
relweights(regres02, col = "orange")
```

## Hacemos el modelo de prediccion
### Con una muestra de diez observaciones
```{r echo=FALSE}
set.seed(1234)

nrow(mData)
n<- 10 

#Elegimos el tamaño muestral y luego los seleccionamos.
ind <- sample(1:nrow(mData),n, replace=FALSE)
NBA_obj<-mData[ind,]
NBA_obj
```

## Hacemos la prediccion, y sacamos el salario de cada jugador
```{r echo=FALSE}
predic<-predict(regres02,newdata=NBA_obj)
predic
```

